---
title: "California County Hospital Data"
author: "Tyler Chia and Joanne Kim"
output:
  html_document:
    theme: darkly
---

```{r, message=FALSE, warning=FALSE, echo = F}
library(tidyverse)
library(knitr)
library(readxl)
library(zoo)
```

For this project we will use multiple historic, county level datasets which is stored as updating CSV's at these URLs:

Hospital Dataset:
```{r, eval = FALSE}
https://hifld-geoplatform.opendata.arcgis.com/datasets/6ac5e325468c4cb9b905f1728d6fbf0f_0
```

Demographic Dataset:
```{r, eval = FALSE}
https://www.census.gov/data/datasets/time-series/demo/popest/2010s-counties-detail.html
```

COVID-19 Dataset:
```{r, eval = FALSE}
https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv
```

The code for this project can be found in my repository [here](https://github.com/TylerChia/hospitals-project).

*****

## Purpose: Analyze various datasets in order to observe if there is any any correlation between race or poverty and the construction of hopsitals within counties of California.

```{r}
hospitals <- read_csv("~/github/hospitals-project/data/Hospitals.csv")
```

```{r}
library(sf)
hos_sf = st_as_sf(hospitals, coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

ca = USAboundaries::us_states() %>% 
  st_transform(4326) %>% 
  filter(stusps == "CA")
ca_counties = USAboundaries::us_counties() %>% 
  st_transform(4326) %>% 
  filter(state_abbr == "CA")
ca_hospitals <- hos_sf %>% 
  filter(STATE == "CA") %>% 
  filter(STATUS == 'OPEN')

ggplot() +
  geom_sf(data = ca, color = "blue") +
  geom_sf(data= ca_counties, color = "blue") +
  geom_sf(data = ca_hospitals, color = "red") +
  coord_sf(datum = st_crs(ca_hospitals)) +
  labs(title = "Location of Hospitals Throughout California",
       x = "Longitude",
       y = "Latitude",
       caption = "Based on Hospitals Dataset")
```

```{r}
num_hospitals <- hospitals %>% 
  filter(STATE == 'CA') %>% 
  filter(STATUS == "OPEN") %>% 
  group_by(COUNTY) %>% 
  summarize(Hospitals = n())

knitr::kable(num_hospitals, caption = "Number of Hospitals Per California County", col.names = c("County", "Number of Hospitals"))
```



